---
title: "Tran_Riddle_Historical Embeddings"
output: github_document
date: December 3, 2024
author: Nela Riddle
---
  
```{r loadPackages, echo = FALSE, message=FALSE, warning=FALSE}
########################################################

## What are the historical patterns of social group representations?
## 14 social groups, 200 years of Google Books Text

########################################################

## Set up R workspace and packages ----
## Set WD to source file location (python codes, stored data, scripts, etc. can all be shared to one folder for ease of use)
# setwd("**")

# load packages
if (!require("corrplot")) {install.packages("corrplot", dependencies = TRUE);require(corrplot)}
if (!require("reticulate")) {install.packages("reticulate", dependencies = TRUE);require(reticulate)}
if (!require("lsa")) {install.packages("lsa", dependencies = TRUE);require(lsa)}
if (!require("dplyr")) {install.packages("dplyr", dependencies = TRUE);require(dplyr)}
if (!require("devtools")) {install.packages("devtools", dependencies = TRUE);require(devtools)}
# if (!require("sweater")) {devtools::install_github("chainsawriot/sweater");require(sweater)} 
if (!require("sweater")) {install.packages("sweater");require(sweater)} 
# contains convenience functions for embeddings analyses 
# see: https://rdrr.io/github/chainsawriot/sweater/f/README.md
```

First, load in the pre-written group and word lists to be used in analyses:
```{r loadWords, include=TRUE, warning=FALSE}
## Load in data ----
## Set WD to word stimuli
setwd("wordstim")

# Function to read and process lists
read_list <- function(file, col_name) {
  list_data <- read.delim(file, header = FALSE)
  colnames(list_data) <- col_name
  return(as.vector(list_data[[col_name]]))
}

# Specific groups (men, women)
groupwrds <- read.csv("groupstimlists.csv", header = FALSE)
groupwrds <- as.data.frame(t(groupwrds))
colnames(groupwrds) <- as.character(groupwrds[1, ])
groupwrds <- groupwrds[-1, ]

# Read lists using the function
agentic <- read_list("agentic.txt", "agentic")
communal <- read_list("communal.txt", "communal")
trait <- read_list("traitlist.txt", "trait")
job <- read_list("joblist.txt", "job")
fruit <- read_list("fruit.txt", "fruit")
noun <- read_list("nouns.txt", "noun")
common <- read_list("common.txt", "common")
```
The agentic and communal lists were borrowed from https://onlinelibrary.wiley.com/doi/10.1002/ejsp.2561; here are some examples
```{r wordExamples}

head(agentic)
head(communal)

```

The group word lists were taken from https://pubmed.ncbi.nlm.nih.gov/35787033/, as well as the trait list:
```{r groupExamples}
head(groupwrds$men)
head(groupwrds$women)
head(trait)
```
The job titles were scraped off this site: https://spotterful.com/blog/job-description-template/job-titles-list-a-z, and expanded through nearest neighbors
```{r jobExamples, echo=FALSE, message=FALSE}
head(job)
```


```{r loadVectors, echo=FALSE, message=FALSE}
# Check unavailable words by decade ----
if (!exists("wordvecs.dat", envir = .GlobalEnv)) {
  load("engall/wordvecsdata_engall.RData", envir = .GlobalEnv)
}
unavwords <- lapply(wordvecs.dat, function(x)
  rownames(x)[which(x$V1 == 0)])
n_avwords <- sapply(wordvecs.dat, function(x)
  length(x$V1) - length(rownames(x)[which(x$V1 == 0)]))

if (!exists("wordvecs.dat_coha", envir = .GlobalEnv)) {
  load("coha/wordvecsdata_coha.RData", envir = .GlobalEnv)
}
unavwords_coha <- lapply(wordvecs.dat_coha, function(x)
  rownames(x)[which(x$V1 == 0)])
n_avwords_coha <- sapply(wordvecs.dat_coha, function(x)
  length(x$V1) - length(rownames(x)[which(x$V1 == 0)])
)
```
The workhorse function; it iterates over each decade, computing the MAC score between each word and each group, then finds the Pearson correlation of the resulting lists (demonstrated visually later)
```{r importantFunction, message=FALSE}
grpwrdassoc_rel <-
  function(group1index,
           group2index,
           wordterms,
           wordvecs.dat = wordvecs.dat,
           unavwords = unavwords,
           corpus) {
    
    start_year <- if (corpus=="coha") 1820 else 1800
    end_year <- if (corpus=="coha") 2010 else 1990
    
    # Create lists of the group's available words
    availwrds_decade_group1 <-
      lapply(1:length(wordvecs.dat), function(i) {
        groupwrds[, group1index][!groupwrds[, group1index] %in% unavwords[[i]]]
      })
    availwrds_decade_group2 <-
      lapply(1:length(wordvecs.dat), function(i) {
        groupwrds[, group2index][!groupwrds[, group2index] %in% unavwords[[i]]]
      })
    
    
    # Now compute MAC from available words for each decade
    wordvecs.mat <- list()
    mac_group1_2list <- list()
    cor_group1_2 <- list()
    cor_group1_2ts <- vector()
    for (i in 1:length(wordvecs.dat)) {
      wordvecs.mat[[i]] <- as.matrix(wordvecs.dat[[i]])
      mac_group1_2list[[i]] <-
        data.frame(
          grp1ef = mac(wordvecs.mat[[i]], S = wordterms, A = availwrds_decade_group1[[i]])$P,
          grp2ef = mac(wordvecs.mat[[i]], S = wordterms, A = availwrds_decade_group2[[i]])$P,
          trait = names(
            mac(wordvecs.mat[[i]], S = wordterms, A = availwrds_decade_group1[[i]])$P
          )
        )
      cor_group1_2[[i]] <-
        cor.test(mac_group1_2list[[i]]$grp2ef, mac_group1_2list[[i]]$grp1ef)
      cor_group1_2ts[i] <- cor_group1_2[[i]]$estimate
      cor_group1_2ts <-
        ts(
          cor_group1_2ts,
          start = start_year,
          end = end_year,
          frequency = 1 / 10
        )
      print(i)
    }
    output_rel <- list(mac_group1_2list,
                       cor_group1_2ts)
    return(output_rel)
  }
```
Demonstrating the mac function in fact works
```{r sanityCheck1}
print(mac(
  as.matrix(wordvecs.dat[[20]]),
  S = c("cat", "puppy", "canine", "happy", "weird", "car"),
  A = c("dog")
)$P)
```
```{r sanityCheck2}
print(mac(
  as.matrix(wordvecs.dat[[20]]),
  S = c("god", "bible", "lord", "banana", "french", "blond"),
  A = c("jesus")
)$P)
```

```{r helperFunctions, echo = FALSE, message = FALSE}
generate_data <- function(group1index,
                          group2index,
                          wordterms,
                          corpus) {
  file_path <- paste0("outputs/",
                      paste(group1index, group2index, wordterms, corpus, sep = "_"),
                      ".RData")
  if (file.exists(file_path)) {
    # Load the object from the .RData file into a temporary environment
    temp_env <- new.env()
    load(file_path, envir = temp_env)  # Load the object
    output <- temp_env$output  # Access the loaded object
  } else {
    if (corpus == "coha") {
      wordvecs.dat_selected <- wordvecs.dat_coha
      unavwords_selected <- unavwords_coha
    } else {
      wordvecs.dat_selected <- wordvecs.dat
      unavwords_selected <- unavwords
    }
    output <- grpwrdassoc_rel(
      group1index = group1index,
      group2index = group2index,
      wordterms = get(wordterms),
      wordvecs.dat = wordvecs.dat_selected,
      unavwords = unavwords_selected,
      corpus = corpus
    )
    
    if (!dir.exists("outputs")) {
      dir.create("outputs")
    }
    save(output, file = file_path)
  }
  
  return(output)
}
get_data_internal <- function(group1index,
                              group2index,
                              wordterms,
                              corpus,
                              decade = NULL) {
  data <- generate_data(
    group1index = group1index,
    group2index = group2index,
    wordterms = wordterms,
    corpus = corpus
  )
  if (is.null(decade)) {
    output <- data[[2]]
  } else {
    if (corpus == "coha") {
      output <- data[[1]][[(decade - 1810) / 10]]
    } else{
      output <- data[[1]][[(decade - 1790) / 10]]
    }
    attr(output, "decade") <- decade
  }
  attr(output, "group1index") <- group1index
  attr(output, "group2index") <- group2index
  attr(output, "wordterms") <- wordterms
  attr(output, "corpus") <- corpus
  
  return(output)
}
get_ts <- function(group1index,
                   group2index,
                   wordterms,
                   corpus) {
  get_data_internal(group1index, group2index, wordterms, corpus)
}
get_decade <-
  function(group1index,
           group2index,
           wordterms,
           corpus,
           decade) {
    get_data_internal(group1index, group2index, wordterms, corpus, decade = decade)
  }
ts1 <- get_ts("men", "women", "agentic", "engall")
ts2 <- get_ts("men", "women", "communal", "coha")
ts3 <- get_ts("men", "women", "noun", "coha")
ts4 <- get_ts("nonhuman", "women", "noun", "coha")



decade1 <- get_decade("men", "women", "agentic", "engall", 1800)
decade2 <- get_decade("men", "women", "agentic", "coha", 1820)
decade2 <- get_decade("men", "women", "noun", "coha", 2010)


ts4

decade2
```

```{r makePlotsRevised, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)

plot_one_ts <-
  function(ts_data) {
    ts_df <- data.frame(Year = seq(from = 1800, to = 1990, by = 10),
                        Value = as.vector(ts_data))
    group1 <- attr(ts_data, "group1index")
    group2 <- attr(ts_data, "group2index")
    word_term <- attr(ts_data, "wordterms")
    corpus <- attr(ts_data, "corpus")
    title <-
      paste(group1, "vs", group2, ",", word_term, "(", corpus, ")")
    p <- ggplot(ts_df, aes(x = Year, y = Value)) +
      geom_line(color = "blue") +
      labs(title = title, x = "Year", y = "Value") +
      theme_minimal()
    
    return(p)
  }

plot_multiple_ts <- function(ts_list) {
  # Create an empty data frame to store all the time series data
  combined_df <- data.frame()
  
  # Loop through each time series in the list and prepare the data for plotting
  for (i in seq_along(ts_list)) {
    ts_data <- ts_list[[i]]
    ts_df <- data.frame(
      Year = seq(from = 1800, to = 1990, by = 10),
      Value = as.vector(ts_data),
      group1 = attr(ts_data, "group1index"),
      group2 = attr(ts_data, "group2index"),
      word_term = attr(ts_data, "wordterms"),
      corpus = attr(ts_data, "corpus")
    )
    combined_df <- rbind(combined_df, ts_df)
  }
  
  # Create the plot with facets by corpus
  p <-
    ggplot(combined_df, aes(
      x = Year,
      y = Value,
      color = interaction(group1, group2, word_term)
    )) +
    geom_line() +
    labs(title = "Multiple Time Series Plot", x = "Year", y = "Similarity Coefficient") +
    facet_wrap( ~ corpus, scales = "fixed") +
    theme_minimal() + guides(color = guide_legend(title = NULL))
  
  return(p)
}

plot_one_decade <- function(decade_data) {
  # Extract attributes
  group1 <- attr(decade_data, "group1index")
  group2 <- attr(decade_data, "group2index")
  word_term <- attr(decade_data, "wordterms")
  corpus <- attr(decade_data, "corpus")
  decade <- attr(decade_data, "decade")
  
  # Calculate Pearson correlation
  pearson_corr <- cor(decade_data$grp1ef, decade_data$grp2ef, use = "complete.obs")
  
  # Create title
  title <- paste0(
    group1, " vs ", group2, ", ", word_term, 
    " (", corpus, ", ", decade, ", r = ", round(pearson_corr, 2), ")"
  )
  
  # Create plot
  p <- ggplot(decade_data, aes(x = grp1ef, y = grp2ef)) +
    geom_point(color = "blue") + geom_smooth(method="lm")+
    geom_text(aes(label = trait), vjust = -0.5, size = 3) +
    labs(title = title, x = group1, y = group2) +
    theme_minimal()
  
  return(p)
}

```

```{r plottingDecade, warning=FALSE}

plot_one_decade(get_decade("men", "women", "agentic", "coha", 1820))
plot_one_decade(get_decade("men", "women", "communal", "engall", 1990))
plot_one_decade(get_decade("men", "women", "common", "engall", 1800))


head(get_decade("men", "women", "agentic", "engall", 1990))
```
```{r propMissing}
# Calculate the number of missing group words by decade
groupmiss_coha <-
  as.data.frame(sapply(1:dim(groupwrds)[2], function(j) {
    sapply(1:length(wordvecs.dat_coha), function(i) {
      sum(groupwrds[, j] %in% unavwords_coha[[i]])
    })
  }))
colnames(groupmiss_coha) <- colnames(groupwrds)
groupmiss_coha[21,] <- colSums(groupwrds != "", na.rm = TRUE)
groupmiss2_coha <-
  as.data.frame(sapply(1:dim(groupmiss_coha)[2], function(j) {
    1 - groupmiss_coha[1:20, j] / groupmiss_coha[21, j]
  }))
colnames(groupmiss2_coha) <- colnames(groupwrds)

# Add the year column
groupmiss2_coha$year <- seq(1820, 2010, by = 10)
groupmiss2_coha


```
```{r plotMissing}
# Create the plot
ggplot(groupmiss2_coha, aes(x = year)) +
  geom_line(aes(y = men, color = "Men")) +
  geom_line(aes(y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "red")) +
  labs(title = "prop. group words available over time",
       x = "Years",
       y = "Values",
       color = "Legend") +
  theme_minimal()
```

```{r outliersMissing}
find_outliers <- function(column, year) {
  is_outlier <-
    abs(column - mean(column, na.rm = TRUE)) > 3 * sd(column, na.rm = TRUE)
  data.frame(Decade = year[is_outlier], Value = column[is_outlier])
}

# Find outliers for men and women
outliers_men <-
  find_outliers(groupmiss2_coha$men, groupmiss2_coha$year)
outliers_women <-
  find_outliers(groupmiss2_coha$women, groupmiss2_coha$year)

# Print results
outliers_men
outliers_women
```


```{r getMagnitudes}
# Initialize an empty data frame to store results
results_df <- data.frame()

# Define the parameters to iterate over
categories <- c("agentic", "communal")
corpora <- c("coha", "engall")
years <- seq(1800, 2010, by = 10)

# Loop through each year
for (year in years) {
  # Create a temporary vector to store the values for the current year
  temp_row <- c(year)
  
  # Loop through each combination of trait and corpus
  for (category in categories) {
    for (corpus in corpora) {
      if ((year >= 1820 & corpus == "coha") | (year <= 1990 & corpus == "engall")) {
        df <- get_decade("men", "women", category, corpus, year)
        mean_men <- mean(df$grp1ef, na.rm = TRUE)
        mean_women <- mean(df$grp2ef, na.rm = TRUE)
        temp_row <- c(temp_row, mean_men, mean_women)
      } else{
        temp_row <- c(temp_row, NaN, NaN)
      }
    }
  }
  results_df <- rbind(results_df, temp_row)
}

# Set the column names for the results data frame
colnames(results_df) <- c(
  "year",
  "agentic_coha_men", "agentic_coha_women",
  "agentic_engall_men", "agentic_engall_women",
  "communal_coha_men", "communal_coha_women",
  "communal_engall_men", "communal_engall_women"
)

# Display the results
generate_trait_corpus_plot <- function(data, trait, corpus) {
  # Dynamically generate column names
  y_men <- paste0(trait, "_", corpus, "_men")
  y_women <- paste0(trait, "_", corpus, "_women")
  
  # Generate title
  title <- paste(trait, corpus, sep = ", ")
  
  # Create the plot
  ggplot(data, aes(x = year)) +
    geom_line(aes_string(y = y_men, color = "'Men'")) +
    geom_line(aes_string(y = y_women, color = "'Women'")) +
    labs(
      title = title,
      x = "Year",
      y = "Mean Values",
      color = "Group"
    ) +
    theme_minimal() +
    scale_color_manual(values = c("Men" = "blue", "Women" = "red"))
}

generate_trait_corpus_plot(results_df, "communal", "coha")
generate_trait_corpus_plot(results_df, "agentic", "coha")
generate_trait_corpus_plot(results_df, "communal", "engall")
generate_trait_corpus_plot(results_df, "agentic", "engall")

```

```{r plottingTsHumanNonhuman}
plot_one_ts(get_ts("nonhuman", "women", "agentic", "coha"))
plot_one_ts(get_ts("nonhuman", "men", "communal", "coha"))

human_nonhuman_ts<-list(get_ts("nonhuman", "women", "trait", "coha"),get_ts("nonhuman", "men", "trait", "coha"), get_ts("nonhuman", "women", "trait", "engall"),get_ts("nonhuman", "men", "trait", "engall"),get_ts("men", "women", "trait", "engall"),get_ts("men", "women", "trait", "coha"))
plot_multiple_ts(human_nonhuman_ts)

plot_one_decade(get_decade("nonhuman", "women", "trait", "coha", 2010))

```

```{r plottingTsMenWomen}
plot_one_ts(get_ts("men", "women", "trait", "coha"))
plot_one_ts(get_ts("men", "women", "job", "coha"))

men_women_trait_job_ts<-list(get_ts("men", "women", "trait", "coha"), get_ts("men", "women", "job", "coha"), get_ts("men", "women", "trait", "engall"),get_ts("men", "women", "job", "engall"))
plot_multiple_ts(men_women_trait_job_ts)

men_women_trait_job_ts<-list(get_ts("men", "women", "agentic", "coha"),get_ts("men", "women", "communal", "coha"), get_ts("men", "women", "agentic", "engall"),get_ts("men", "women", "communal", "engall"),get_ts("men", "women", "common", "engall"),get_ts("men", "women", "common", "coha"))
plot_multiple_ts(men_women_trait_job_ts)

# plot_one_decade(get_decade("nonhuman", "women", "trait", "engall", 1990))

plot_one_decade(get_decade("nonhuman", "women", "trait", "coha", 1850))


```